---
title: "Class11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” >
“by Experimental Method and Molecular Type”). Move this CSV file into your RStudio project
and determine the percentage of structures solved by X-Ray and Electron Microscopy. 

```{r}
data_summary <- read.csv("Data Export Summary.csv")
ans <- data_summary$Total/sum(data_summary$Total)*100
names(ans) <- data_summary$Experimental.Method
round(ans, 2)
```

Also can you determine what proportion of structures are protein?
```{r}
round(sum(data_summary$Proteins)/sum(data_summary$Total)*100, 2)
```


# Working with biomolecular data in r
```{r}
library(bio3d)
pdb <- read.pdb("1hsg")
pdb
```
```{r}
attributes(pdb)
```

```{r}
head(pdb$atom)
```
Print a subset of $atom data for the first two atoms
```{r}
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]
```
Note that individual $atom records can also be accessed like this
```{r}
pdb$atom$elety[1:2]
```
Which allows us to do the following:
```{r}
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab="B-factor")
```

You can also pull up a particular coordinate, for example to determine the res ID of atom 2. 
```{r}
pdb$atom[2, "resid"]
```

# Atom selection

This will save a new file to your working directory.
```{r}
protein <- atom.select(pdb, "protein", value=TRUE)
write.pdb(protein, file="1hsg_protein.pdb")


ligand <- atom.select(pdb, "ligand", value=TRUE)
write.pdb(ligand, file="1hsg_ligand.pdb")
```


# Let's try an NMA

```{r}
pdb <- read.pdb("1hel")
modes <- nma(pdb)
m7 <- mktrj(modes,
 mode=7,
 file="mode_7.pdb")
view(m7, col=vec2color( rmsf(m7) ))
```

