---
title: "Class14"
author: "Alanna Koehler"
date: "11/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(BiocManager)
library(DESeq2)
```

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
head(counts)
```
```{r}
head(metadata)
```

How many genes are in this dataset?
```{r}
nrow(counts)
```
How many experiments are there?
```{r}
ncol(counts)-1
```

Let's make sure the metadata id columnds match the column names of counts
```{r}
colnames(counts)[-1]
```
```{r}
metadata$id
```
```{r}
all(colnames(counts)[-1] == metadata$id) #wrapping in all() gives you a summary
```

#Never do differential expression analysis this way!
Summarize all the control columns into one number (the mean)
```{r}
control <- metadata[metadata[,"dex"]=="control",] #The dex column is treated or control
control.mean <- rowSums( counts[ ,control$id] )/length(control$id)
names(control.mean) <- counts$ensgene #Indicate that the names for control.mean should come from counts column ensgene
```

Now summarize all the treated columns into one number
```{r}
treated <- metadata[metadata[, "dex"]=="treated",] 
treated.mean <- rowSums( counts[ ,treated$id])/length(treated$id)
names(treated.mean) <- counts$ensgene
```

Combine mean counts data for bookkeeping purposes
```{r}
meancounts <- data.frame(control.mean, treated.mean)
head(meancounts)
```

You can also find the means using the following code:
```{r}
control.id <- metadata[metadata$dex=="control",]$id
control.id
head(rowSums(counts[,control.id])/length(control.id))
```

```{r}
plot(meancounts) #You can simply plot meancounts or you can specify x and y as below
plot(meancounts[,1],meancounts[,2], xlab="Control", ylab="Treated")
```
Now log transform the axes
```{r}
plot(meancounts[,1],meancounts[,2], xlab="Control", ylab="Treated", log ="xy")
```

Calculate log2foldchange, add it to meancounts data.frame and inspect the results
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

There are a couple of “weird” results. Namely, the NaN ("not a number") and -Inf (negative infinity) results. The NaN is returned when you divide by zero and try to take the log. The -Inf is returned when you try to take the log of zero. It turns out that there are a lot of genes with zero expression. 

Let’s filter our data to remove these genes.
The arr.ind=TRUE (array indices) argument will clause which() to return both the row and column indices (i.e. positions) where there are TRUE values. In this case this will tell us which genes (rows) and samples (columns) have zero counts.
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE) #which() returns the indices for which a condition is TRUE. When you apply it to a dataframe with arr.ind=TRUE, it will return a dataframe rather than a vector of the number of each element.
head(zero.vals)
```

We are going to ignore any genes that have zero counts in any sample so we just focus on the row answer. Calling unique() will ensure we dont count any row twice if it has zero entries in both samples. 

```{r}
to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

How many genes do we remove?
```{r}
length(to.rm)
```

How many unique genes are we left with?
```{r}
nrow(mycounts)
```

A common threshold used for calling something differentially expressed is a log2(FoldChange) of greater than 2 or less than -2.
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

```{r}
count(up.ind)
sum(up.ind) #You can also use the sum button because TRUE==1.
```

```{r}
count(down.ind)
sum(down.ind)
```

Our mycounts result table so far only contains the Ensembl gene IDs. However, alternative gene names and extra annotation are usually required for informative for interpretation. We can add annotation from a supplied CSV file, such as those available from ENSEMBLE or UCSC. The annotables_grch38.csv annotation table links the unambiguous Ensembl gene ID to other useful annotation like the gene symbol, full gene name, location, Entrez gene ID, etc.
```{r}
anno <- read.csv("annotables_grch38.csv")
head(anno)
```

You can also use a Bioconductor package if you don't have a favorite annotation library.
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
columns(org.Hs.eg.db)
```

We can use the mapIds() function to add individual columns to our results table. We provide the row names of our results table as a key, and specify that keytype=ENSEMBL. The column argument tells the mapIds() function which information we want, and the multiVals argument tells the function what to do if there are multiple possible values for a single input value. Here we ask to just give us back the first one that occurs in the database.
```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts), # Our genenames
                     keytype="ENSEMBL",        # The format of our genenames
                     column="SYMBOL",          # The new format we want to add
                     multiVals="first")
```

```{r}
head(mycounts)
```
Add Entrez and Uniprot
```{r}
mycounts$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")

mycounts$uniprot <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="UNIPROT",
                     keytype="ENSEMBL",
                     multiVals="first")
head(mycounts)
```
```{r}
View( mycounts[up.ind,] )
```



#Exercise in using unique()
```{r}
x <- data.frame(happy=c(1,0,4,3,2,4), sad=c(4,0,0,2,0,4))
x
```
```{r}
x==0
which(x==0, arr.ind=TRUE)
```
```{r}
unique(which(x==0, arr.ind=TRUE)[,1])
```

