---
title: "Class 8 Machine Learning 1"
author: "Alanna Koehler"
date: "10/25/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## K-means example

Generate a simple dataset for clustering: 
```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3)) #Give you 30 points centered at -3 and 30 points centered at 3.
x <- cbind(x=tmp, y=rev(tmp)) #column-wise binds our vector into 2 columns
plot(x)
```

Use the kmeans() function setting k to 2 and nstart=20
```{r}
k <- kmeans(x, centers = 2, nstart = 20)

```
This gives us a list of 9 elements in our global environment. Print the results
```{r}
k
```
The clustering vector tells you in which cluster each element lies.

Inspect the results:
Q. How many points are in each cluster?
```{r}
k$size
```
  
Q. What ‘component’ of your result object details
 - cluster size?
```{r}
k$cluster
```
 
 - cluster assignment/membership?
```{r}
table(k$cluster)
```
 
 - cluster center?
```{r}
k$centers
```
 
Plot x colored by the kmeans cluster assignment and add cluster centers as blue points
```{r}
plot(x, col = k$cluster)
points(k$centers, col = "blue", pch = 15)
```
 
## Hierarchical clustering example
First we need to calculate point (dis)similarity as the Euclidean distance between observations. 
```{r}
dist_matrix <- dist(x)
```
The hclust() function returns a hierarchical clustering model
```{r}
hc <- hclust(d = dist_matrix)
```
The print method is not so useful here
```{r}
hc
```
Plot results in a dendrogram
```{r}
plot(hc)
```
You then have to cut this tree in a position you like. First you can generate an abline:
```{r}
plot(hc)
abline(h=6, col="red")
abline(h=4, col="blue")
```
Cut the tree at a certain height and assign each element to a cluster. 
```{r}
cutree(hc, h=6)
cutree(hc, h=4)
```
You can quantify how many elements are in each cut group by making a table.
```{r}
grps <- cutree(hc, h=6)
table(grps)
```
Once you have defined the groups, you can then color your clusters based on those groups.
```{r}
plot(x, col = grps)
```
You can also cut the tree to yield a given k groups/clusters
```{r}
cutree(hc, k=2)
```

## Using different hierarchical clustering methods
Step 1. Generate some example data for clustering
```{r}
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```
Step 2. Plot the data without clustering
```{r}
plot(x)
```
Step 3. Generate colors for known clusters (just so we can compare to hclust results).
```{r}
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

Answer questions
Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters. How does this compare to your known col groups. 
```{r}
hc2 <- hclust(dist(x))
plot(hc2)
abline(h=2.5, col = "red")
abline(h=2, col = "blue")
```

```{r}
grps2 <- cutree(hc2, k=2)
table(grps2)
```
```{r}
plot(x, col = grps2)
```

```{r}
table(col, grps2)
```

```{r}
grps3 <- cutree(hc2, k=3)
table(grps3)
```
```{r}
plot(x, col = grps3)
```
```{r}
table(col, grps3)
```


## Principle Component Analysis

Plot example data
```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1)
head(mydata)
```
Note that samples are columns and genes are rows. 

Now define your PCA components and look at the attributes
```{r}
pca <- prcomp(t(mydata), scale=TRUE) #Need to transpose with t() bc prcomp expects samples to be rows and genes to be columns
attributes(pca) 
```
Make a basic PC1 vs PC2 plot.
```{r}
plot(pca$x[,1], pca$x[,2]) 
```
You can take a look at the importance of components using the summary()
```{r}
summary(pca)
```
Now you can see that PC1 captures 92.6% of the variance. 

Calculate variance captured by each PC.
```{r}
pca.var <- pca$sdev^2 
pca.var
```
You can also look at percent variance.
```{r}
pca.var <- pca$sdev^2 #square the standard deviation
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)  #variance divided by total possible sum in percent
pca.var.per
```
Make a scree plot.
```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```

Add color to your PCA plot.
```{r}
colvec <- as.factor( substr( colnames(mydata), 1, 2) ) #substr() takes a substring
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```

 
 