---
title: 'Class 08: UK PCA'
author: "Alanna Koehler"
date: "10/25/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
x <- read.csv("UK_foods.csv")
x <- read.csv("data/UK_foods.csv")
```

Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?
```{r}
dim(x)
```

Fix row names so they are no longer the first column
```{r}
rownames(x) <- x[,1] #This is the suboptimal method.
x <- x[,-1]
head(x)
```
```{r}
dim(x)
```

This method is perhaps better.
```{r}
x <- read.csv("data/UK_foods.csv", row.names=1)
head(x)
```

Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?
If you use the first approach it removes a column each time you call it.

Plot the data:
```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

Q3: Changing what optional argument in the above barplot() function results in the following plot?

```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

Q5: Generating all pairwise plots may help somewhat. Can you make sense of the following code and resulting figure? What does it mean if a given point lies on the diagonal for a given plot?
```{r}
pairs(x, col=rainbow(10), pch=16)
```
Northern Ireland is the outlier. If a given point lies on the diagonal there is a linear relationship between countries. 

Q6. What is the main differences between N. Ireland and the other countries of the UK in terms of this data-set?
The blue data point.

Make a PCA plot.
```{r}
pca <- prcomp( t(x) )
summary(pca)
```

Q7. Complete the code below to generate a plot of PC1 vs PC2. The second line adds text labels over the data points.

```{r}
plot(pca$x[,1:2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```

Q8. Customize your plot so that the colors of the country names match the colors in our UK and Ireland map and table at start of this document.

```{r}
plot(pca$x[,1:2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x), col = c("orange", "red", "blue", "green"))

```





Variable Loadings
```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

