---
title: 'Class 6: R Functions'
author: "Alanna Koehler"
date: "10/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This is an H1 heading
This is my class 6 work. 

## This is an H2 heading

### This is an H3 heading

```{r}
plot(1:10)
```


```{r}
read.csv("test1.txt")
read.table("test2.txt", header = TRUE, sep = "$")
read.table("test3.txt", sep = "")
```

Our first silly
```{r}
add <- function(x, y=1) {
 # Sum the input x and y
 x + y}
```


```{r}
add(c(5,6))
```
 
 If you add three values, you wil get an error because you only specified x and y in the function above. You can, however, use a vector.
 
```{r}
add (c(5,5,NA,7))
```
 NA holds a space in your dataframe and returns NA. 
 
Writing your own function:
You need a “name”, “arguments” and “body”. If you plan on having NA values, you need to indicate na.rm = TRUE to omit those values.
```{r}
rescale <- function(x) {
 rng <-range(x, na.rm = TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}
```
Click the play button to add to your environment. Test with a known value.

```{r}
rescale(1:10)
```


```{r}
rescale(c(1,2,NA,5))
```

Try this function:
```{r}
rescale2 <- function(x, na.rm=TRUE, plot=FALSE) {
 if(na.rm) {
 rng <-range(x, na.rm=na.rm)
 } else {
 rng <-range(x)
 }
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
}
```

Test this function:
```{r}
rescale2(1:10)
```

