---
title: "Class06 Homework"
author: "Alanna Koehler"
date: "10/18/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load library bio3d.
```{r}
library(bio3d)
```

Write a function plot_pdb() that will read a PDB file and trim it to contain just chain A and alpha carbons.
```{r}
plot_pdb <- function(x) {
  trim.pdb(read.pdb(x), chain="A", elety="CA") 
}
```

Execute new function plot_pdb() for each PDB file and assign a new variable name for each.
```{r}
s1 <- plot_pdb("1AKE") #without drug
s2 <- plot_pdb("4AKE") #with drug A
s3 <- plot_pdb("1E4Y") #with drug B
```

Make a line plot of the "atom" and "b" columns from the first variable (s1), along with SSE along the x axis (Residue). The y axis is Bfactor. Now add the other 2 variable sets (s2 and s3) in different colors to the line plot. Generate a color-coded legend in the top right corner and a title at the top.
```{r}
plotb3(s1$atom$b, sse=s1, top = FALSE, typ="l", ylab="Bfactor") #plot the first variable
points(s2$atom$b, typ="l", ylab="Bfactor", col="red") #add a second variable
points(s3$atom$b, typ="l", ylab="Bfactor", col="blue") #add third variable
leg = c("no drug", "drug A", "drug B") #define legend text
legend("topright", legend = leg, fill = c("black", "red", "blue"), bty = "n", inset = 0) #generate a legend
title(main = "Protein Drug Interactions") #generate a title
```
