---
title: "Class07 R Functions and Packages"
author: "Alanna Koehler"
date: "10/23/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Revisiting functions from last class

```{r}
source("http://tinyurl.com/rescale-R")
#These files are read only.
```

Let's try our rescale() function from last class
```{r}
rescale(1:10)
```

```{r}
rescale(c(3, 10, NA, 7))
#Remember we fixed our function so it doesn't propagate NA. If you insert a string, however, you will get an error message. 
```

```{r}
rescale2("barry")
#This new function adds an "if" loop to stop the process if the input is not a numeric AND return a unique error message.
#if( !is.numeric(x) ) {
 #stop("Input x should be numeric", call.=FALSE)
 #}
```

We want to write a function, called both_na(), that counts how many positions in two input vectors, x and y, both have a missing value:
Tip: Search for existing functionality to get us started. Get a working snippet of code that is close to what we want then rewrite and test various inputs (ie. eejit proofing)
```{r}
#First define your variables
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
#Then find a pre-existing function
is.na(x) #determines if x value is na
is.na(y) #determines if y value is na
is.na(x) & is.na(y) #in which cases are both of these things TRUE
```

```{r}
#Refine those pre-existing functions and test
sum(is.na(x) & is.na(y)) #determines how many indices there are where x AND y are both na
```
```{r}
#Define your own function
both_na <- function(x,y) { #assign to a name and define arguments
  sum(is.na(x) & is.na(y)) #write the body
}
#Now test that new function with your original scenario
both_na(x,y)
```
```{r}
#Now test your function with new parameters
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
y3 <- c(1, NA, NA, NA, NA, NA)
both_na(x, y1) #Return data for x and y1
both_na(x, y2) #Return data for x and y2... Gives you a warning that the vectors are not equal length but returns an answer of 3 by recycling x.
both_na(x, y3) #Return data for x and y3...Does NOT give you a warning because y3 is double the length of x. However, because x, the shorter vector, is recycled, the count is 5.
```

```{r}
#Aside on recycling
plot(1:10, col=c("red", "blue", "green")) 
#In this example, we recycle through colors.
```

```{r}
#Now revise to make your function better
both_na2 <- function(x,y) {
   if(length(x) != length(y)) { #Put your stop or warning as early as possible!
    stop("Input vectors should be the same length!") #If you don't want the original error message, use .call=FALSE
  }
  sum(is.na(x) & is.na(y)) 
}

#Retest the revised function. Does it work as expected in both cases?
both_na2(x, y1)
both_na2(x, y3)
```
```{r}
#You can make this even more complex if desired:
both_na3 <- function(x, y) {
 if(length(x) != length(y)) {
 stop("Input x and y should be vectors of the same length")
 }

 na.in.both <- ( is.na(x) & is.na(y) )
 na.number <- sum(na.in.both)
 na.which <- which(na.in.both)
 message("Found ", na.number, " NA's at position(s):",
 paste(na.which, collapse=", ") )

 return( list(number=na.number, which=na.which) )
}

#Test it
both_na3(x, y1)
both_na3(x, y3)
```



##In Class Exercise
Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single alignment score.

```{r}
# student 1
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
# student 2
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
#student 3
student3 <- c(100, NA, NA, NA, NA, NA, NA, NA)
```

```{r}
grade <-function(x){
  x[is.na(x)] <- 0 #Set NA equal to 0
  round(mean(x[-which.min(x)])) #Take the mean except the min value and round.
}
grade(student1)
grade(student2)
grade(student3)
```

Now grade all students in an example class
url <- "https://tinyurl.com/gradeinput"

